% !TeX program = xelatex (if the compiler is set to pdflatex, change it!)

%=====================================================================
\documentclass[11pt,aspectratio=169]{beamer}
\usetheme{metropolis}
\usepackage{fontspec}
\usepackage{appendixnumberbeamer} % for \inserttotalframenumber
\usepackage{graphicx,booktabs,amsmath,amssymb,tikz}
\usepackage{hyperref}
\usepackage{pifont}
\usepackage{xcolor}
\usepackage[backend=biber]{biblatex}

%---------------- Bibliography ----------------
\addbibresource{/Users/lucasdubois/Desktop/LaTeX/MASTER/master.bib}

%--------------Images----------------
\graphicspath{{Images/}}

% ---------------- Fonts ----------------
\setsansfont{Fira Sans}[
  Ligatures=TeX,
  Scale=MatchLowercase,
  UprightFont     = * Book,
  ItalicFont      = * Book Italic,
  BoldFont        = * SemiBold,
  BoldItalicFont  = * SemiBold Italic
]
\setmonofont{Fira Mono}[Scale=MatchLowercase]

\title{\textsc{Dividend Taxation and Top-Income Inequality in Brazil}}
\author{Lucas Dubois}
\institute{HEC Liège}
\date{\today}

% ---------------- Executive Palette ----------------
\definecolor{primaryred}{RGB}{127,20,22}
\definecolor{midnight}{RGB}{5,26,40}
\definecolor{steelblue}{RGB}{171,193,223}
\definecolor{offwhite}{RGB}{230,230,220}
\definecolor{darkgreen}{RGB}{34,139,34}

\setbeamercolor{background canvas}{bg=offwhite}
\setbeamercolor{normal text}{fg=midnight}
\setbeamercolor{title}{fg=midnight}
\setbeamercolor{subtitle}{fg=midnight!80}
\setbeamercolor{author}{fg=midnight!75}
\setbeamercolor{date}{fg=midnight!75}
\setbeamercolor{frametitle}{fg=steelblue,bg=midnight}

% Progress bar colors (used on section pages)
\setbeamercolor{progress bar}{fg=primaryred,bg=midnight!30}
\setbeamercolor{progress bar in section page}{fg=primaryred,bg=midnight!30}
\setbeamercolor{title separator}{fg=primaryred,bg=primaryred!25}
% Blocks very light (demo-style)
\setbeamercolor{block title}{fg=midnight,bg=steelblue!12}
\setbeamercolor{block body}{fg=midnight,bg=steelblue!6}

% ---------------- Metropolis options ----------------
% Show ONLY a section page between sections with a progress bar.
% No per-frame progress under the frametitle.
\metroset{
  numbering=fraction,
  progressbar=none,          % no frametitle progress bar
  sectionpage=progressbar,   % section pages with progress bar (single line)
  block=transparent,
  titleformat title=regular,
  subsectionpage=none
}

% Make the section-page progress bar thicker
\makeatletter
\setlength{\metropolis@progressonsectionpage@linewidth}{1.8pt} % thicker than default
\makeatother

\hypersetup{colorlinks=true,linkcolor=primaryred,urlcolor=primaryred}

\begin{document}

% ========== Title page ==========
\maketitle

% ========== Outline ==========
\begin{frame}{\textsc{Table of Contents}}
  \vspace{1.0ex}
  \tableofcontents
\end{frame}

% ================== Content ==================
\section{\textsc{Context:}}
\begin{frame}{\textsc{Brazilian Taxation and Inequality:}}
Since the colonial period, Brazil has been characterized by high levels of income inequality:
  \begin{itemize}
    \item \textcolor{primaryred}{Gini}: 0.51 (\textcolor{steelblue}{5th highest globally}).
    \item \textcolor{primaryred}{Wealth concentration:} Top 1\% owns \textcolor{steelblue}{48.4\%} of total wealth (\textcolor{steelblue}{1st globally}) (2022).
    \item \textcolor{primaryred}{Regressive tax system:} \textcolor{steelblue}{42.5\%} Average tax rate of  vs. \textcolor{steelblue}{20.6\%} for million-dollar earners (\cite{zucmanpalomo}).
  \end{itemize}
\textsc{\color{steelblue}But in 1995...} \\
\end{frame}

\begin{frame}{\textsc{Definition:}}
    What is {\color{steelblue}\textbf{Machine Learning (ML)}}?
    \begin{enumerate}
        \item {\color{primaryred}High-dimensional models.}
        \begin{itemize}
            \item[\textcolor{darkgreen}{\ding{51}}] Flexibility.
            \item[\textcolor{red}{\ding{55}}] Overfitting.
        \end{itemize}
        \item {\color{primaryred}“Regularization” methods for model selection.}
            \begin{itemize}
                \item[\textcolor{darkgreen}{\ding{51}}] Fix Overfitting.
            \end{itemize}
        \item {\color{primaryred}Efficient algorithms for searching model specifications.}
            \begin{itemize}
                \item[\textcolor{darkgreen}{\ding{51}}] Fix non-linearities.
            \end{itemize}
    \end{enumerate}
\end{frame}


\begin{frame}{\textsc{Dataset and Metrics:}}
    \begin{itemize}
        \item {\color{primaryred}\textbf{Universe:}} Monthly U.S. equities (CRSP/Compustat)
        \item {\color{primaryred}\textbf{Sample:}} {\color{steelblue}\textbf{30,000 individual stocks}}.
        \item {\color{primaryred}\textbf{Period:}} {\color{steelblue}\textbf{60 years}} (1957–2016)
        \item {\color{primaryred}\textbf{Predictors:}} {\textbf{94 firm-level characteristics}} from the "factor zoo" + {\textbf{74 industry-sector dummies}} + interaction terms.
        \begin{itemize}
            \item {\color{steelblue} \textbf{900+ predictors}} in total.
        \end{itemize}
    \end{itemize}
\end{frame}

\section{\textsc{Results:}}
\section{\textsc{Wrap-up}}
\begin{frame}{Conclusion}
  \begin{itemize}
    \item Using a single, thicker Metropolis section progress bar.
    \item Clean frametitles with no extra progress line on content slides.
    \item Palette and typography unchanged.
  \end{itemize}
\end{frame}

\section{\textsc{Hello}}
\begin{frame}{Conclusion}
  \begin{itemize}
    \item Using a single, thicker Metropolis section progress bar.
    \item Clean frametitles with no extra progress line on content slides.
    \item Palette and typography unchanged.
  \end{itemize}
\end{frame}


\section{\textsc{Hello again}}
\begin{frame}{Conclusion}
  \begin{itemize}
    \item Using a single, thicker Metropolis section progress bar.
    \item Clean frametitles with no extra progress line on content slides.
    \item Palette and typography unchanged.
  \end{itemize}
\end{frame}

\begin{frame}{\textsc{References:}}
   \printbibliography[heading=none]
\end{frame}

{
\setbeamercolor{background canvas}{bg=midnight}
\begin{frame}[plain]
  \centering
  {\color{primaryred}\fontsize{150}{150}\selectfont ?}
\end{frame}
}
\end{document}
